\documentclass{article}

\usepackage{amsmath}
\usepackage{amssymb}

\newcommand{\R}{\ensuremath{\mathbb{R}}}

\title{Complexity: Approximation Algorithms}
\author{Kunwar Shaanjeet Singh Grover}
\date{4 Aug 2020}

\begin{document}

\pagenumbering{gobble}
\maketitle
\newpage
\pagenumbering{arabic}

\section{Introduction}

What happens when you discover a problem is NP Complete or NP Hard? You can try
to think of a polynomial time algorithm and try to prove that it gives an answer
within some factor of the correct answer. That's what we are going to do today.\\

We will take a bunch of NP Complete problems and try to come up with some
interesting hueristics and try to prove that they give answer within some 
constant factor.

\begin{enumerate}
    
    \item  
    \item Vertex Cover
    \item Set cover
    \item Partition

\end{enumerate}

\section{Approximation Algos and Schemes}

An algorithm for a problem of size n has an approximation ratio $\rho(n)$ if 
for any input, algo produces a solution with cost $C$ s.t max $\left(\frac{C}{Copt}, \frac{Copt}{C}\right) \le \rho(n)$ \\

An approximation scheme takes as input $\epsilon < 0$ and for any fixed $\epsilon$, 
the scheme is a $(1 + \epsilon)$ - approximation algorithm. \\

$O(n^{\frac{2}{\epsilon}})$ running time - Polynomial approx time scheme (PTAS)\\
Polynomial in $n$ but not necessarily in $\epsilon$ \\

Fully PTAS: poly in $n$ and $\frac{1}{\epsilon}$ \\
Example: $O\left(\frac{n}{\epsilon^{2}}\right)$

\subsection{Vertex Cover}

Given: Undirected Graph $G(V, E)$\\

(Vertex Cover 1)\\

Problem: Find a subset $V' subset V$ s.t. if $(u, v)$ is an edge of $G$, then either
$u \in V'$ or $v \in V'$ or both. Find $V'$ so $|V'|$ is minimum.\\

(Vertex Cover 2)\\

Optimal Solution: $k! = n$

Algo could pick all bottom vertices. Solution: 

$$
    k!\left(\frac{1}{k} + \frac{1}{k - 1} + \dots + 1\right)
    \approx k!\log{k}
$$

\subsubsection{Approx Vertex Cover}

\begin{align*}
C \leftarrow \phi \\
E' \leftarrow E
\end{align*}

Pick $(u, v) \in E$ arbitarily

$$
    C \leftarrow C \cup \{u\} \cup \{v\}
$$

Delete from all edges incedent on $u$ or $v$.\\
Return $C$ \\

Approx-Vertex Cover is a 2-approximation algo (Is within a factor of 2 from
the optimal answer).\\

Let $A$ denote the edges that are picked. $C = 2|A|$ vertices are picked. Show that
$Copt \ge |A|$. \\

We need to cover every edge in $A$ since they
are picked this way. Thus, we need to pick a different vertex from each edge in
$A$. Thus $Copt \ge |A|$.

\subsection{Set Cover}

Given: A set $X$ and a family of (possibly overlapping) subsets $S_1, S_2,
\dots, S_m \subseteq X$. s.t.

$$
    \cup_{i = 1}^{m} S_{i} = X
$$

Find $C \subseteq \{1, 2, \dots, m\}$ s.t.

$$
    \cup_{i \in C} = X
$$

while minimizing $|C|$.

(Figure Set Cover)\\

Optimum Solution: $S_{3}, S_{4}, S_{5}$.\\

\subsubsection{Approximate Set Cover}

This is a $(ln(n) + 1)$ approx algo.\\

Pick largest $S_{i}$, remove all elements of $S_{i}$ from $X$ and other $S_{j}$
and repeat.\\

\textbf{Proof}: Assume there is a cover $Copt, |Copt| = t$. Let $X_{k}$ be a set 
of element in iteration $k (X_{0} = X)$. $\forall k, X_{k}$ can be covered by
$t$ sets.

\end{document}
